---
title: "Assignment 4"
author: "Chao XIA"
date: "04/11/2019"
output:
  html_document:
    df_print: paged
---

# Part 1

## Description

1. Use the Urn model starting with 2 red balls of weight 1, and one black (mutation) ball of weight w;

2. Draw balls until you have 10 non-black balls;

3. If all non-black balls are the same color at the end, what is the posterior distribution of the weight of the black ball?

4. Use a Uniform[0,20] prior for the weight of the black ball

## The Urn function

```{r}
Urn_G <- function(finalballamount, starting_ball, black_weight){
  
  ball <- starting_ball # define the starting cituation
  
  while(length(ball) < finalballamount){
    
    pick <- sample(ball, 1, prob  = c(rep(1, length(ball)-1), black_weight)) # pick one ball from the urn
    
    if (pick == "0"){ # 0 is black
      
      ball <- ball[-which(ball == pick)] # pick the black ball
      
      pick2 <- sample(1:length(ball), 1) # pick a ball from the remaining balls
      
      ball[pick2] <- as.character(length(ball) + 1) # change the ball's color (since each iteration the length of ball will increase or remain the same, we'll never repeat the ball's color)
      
      ball <- c(ball, pick) # put back the black ball
      
    }else{
      ball <- ball[-length(ball)]
      
      ball <- c(ball, pick, "0") # else just add a ball with the picked color and keep the black ball at last of the row
      
    }
  }
  
  return(ball)
}
```

## Some Necessary Parameter

```{r}
NoReps <- 10000 # number of iterations
 
Howmanycolorneeded <- 1 # number of color needed

Maxweight <- 20 # Max weight of blackball

Noofballs <- 11 # Number of balls need including blackball
```

## Code and result

```{r}
Acceptedweights <- rep(-9, NoReps)

Starttime <- Sys.time()

for (i in 1:NoReps) {
  HowManyColorsObserved<- -9
  while (HowManyColorsObserved != Howmanycolorneeded) {
    # Sample a weight, ThisWeight, for the mutation ball, from Unif(0,MaxWeight). Then...
    ThisWeight <- sample(Maxweight,1)
    Final_state <- Urn_G(Noofballs, c(1,1,0), ThisWeight) # simulate the urn using this weight
    HowManyColorsObserved <- length(table(Final_state)) - 1
  }
  Acceptedweights[i]<-ThisWeight
}

Finishtime <- Sys.time()

hist(Acceptedweights, breaks = Maxweight, main = paste("Histogram of Accepted Weights", "\n", "Time =", round((Finishtime - Starttime),2), "minutes"))
```

# Part 2

## Description

1. Repeat the Urn model assignment to find the posterior distribution of the weight of the black ball given that we observe just one color in an urn containing 5 balls;

2. This time, sample $\theta'$ from $\xi$ ~ exponential($\lambda$), rather than a Uniform distribution;

3. If you accept that value of $\theta'$ (because the simulated urn has just one color in it), then add a mass of $\frac{\pi(\theta')}{\xi(\theta')}$  to the posterior distribution (of 􀀉􀀁􀀁􀀂whereas we used to add a mass of 1);

4. If we truncate the exponential at 20 (say), this means:

   $\pi(\theta') = \frac{1}{20}$
   
   $\xi(\theta') = \frac{\lambda e^{-\lambda \theta'}}{1-e^{-20 \lambda}}$
   
5. Compare efficiency to a rejection method that just samples directly from a Uniform[0,20] distribution. (i.e. how many simulations do you need to generate 10000 acceptances; and how smooth does the histogram look?)

## Code and Result

```{r}
library(plotrix)
Acceptedweights <- rep(-9, NoReps)
mass <- rep(-9, NoReps)

lambda <- 2

Starttime <- Sys.time()

for (i in 1:NoReps) {
  HowManyColorsObserved<- -9
  while (HowManyColorsObserved != Howmanycolorneeded) {
    theta <- rexp(1, rate = lambda)
    while (theta > 20){
    theta <- rexp(1, rate = lambda)
    }
    Final_state <- Urn_G(Noofballs, c(1,1,0), theta) # simulate the urn using this weight
    HowManyColorsObserved <- length(table(Final_state)) - 1
  }
  Acceptedweights[i] <- theta
  mass[i] <- (1/Maxweight/(lambda*exp(-lambda*theta)/(1-exp(-Maxweight*lambda))))
}

Finishtime <- Sys.time()

weighted.hist(Acceptedweights, mass, main = paste("Histogram of Accepted Weights via Importance sampling", "\n", "Time =", round((Finishtime - Starttime),2), "Seconds"), breaks = 100)
```

The importance sampling using obviously less time than sampling from uniform distribition but the histogram is worse in importance sampling. Importance sampling has more noise. 

## Lambda

Using different Lambad to see the influence of $\lambda$

```{r}
lambda_can <- c(0.5, 1, 2, 4, 8)

par(mfrow = c(2,3))

for (i in lambda_can) {
  
  Acceptedweights <- rep(-9, NoReps)
  mass <- rep(-9, NoReps)
  
  for (j in 1:NoReps) {
    HowManyColorsObserved<- -9
    while (HowManyColorsObserved != Howmanycolorneeded) {
      theta <- rexp(1, rate = lambda)
      while (theta > 20){
      theta <- rexp(1, rate = lambda)
      }
      Final_state <- Urn_G(Noofballs, c(1,1,0), theta)
      HowManyColorsObserved <- length(table(Final_state)) - 1
    }
    Acceptedweights[j] <- theta
    mass[j] <- (1/Maxweight/(lambda*exp(-lambda*theta)/(1-exp(-Maxweight*lambda))))
  }
  
  weighted.hist(Acceptedweights, mass, main = paste("lambda =", i), breaks = 100)
}
```

The influence of $\lambda$ is a little different with larger weight. Some large weights have high importance.


# Part 3

## Description

1. Many tissues contain 2 (or more) cell-types;

2. Investigator wants a way of testing whether each of 2 cell types in a given tissue is homogeneously distributed (in space);

3. Your job, is to come up with such a test, using Monte Carlo methods.

## Threshold

In my opinion, if a grid is homogeneouly distributed, then the coordinate of the tissues with 1 should be randomly distributed, we can use the distribution of spot with 1 to judge if the grid is randomly distributed. I'm using variance of x and y to judge the distribution.

## Some functions and reading the data

```{r}
Grid1 <- read.table("D:/study/MasterinUSC/pm520/week10/grid1.txt", header = F)

Grid2 <- read.table("D:/study/MasterinUSC/pm520/week10/grid2.txt", header = F)

Grid3 <- read.table("D:/study/MasterinUSC/pm520/week10/grid3.txt", header = F)

# generate new grid

Permutate_grid <- function(x){
  nrow_grid <- nrow(x)
  ncol_grid <- ncol(x)
  x <- as.matrix(x)
  x <- as.vector(x)
  new_grid <- sample(x)
  new_grid <- matrix(data = new_grid, nrow = 25, ncol = 25)
  new_grid <- as.data.frame(new_grid)
  return(new_grid)
}

# judgement

Coordinate <- function(x){
  nrow_grid <- nrow(x)
  ncol_grid <- ncol(x)
  Distance <- numeric()
  for(i in 1:nrow_grid){
    for(j in 1:ncol_grid){
      if(x[i,j] == 1){
        Distance <- rbind(Distance, c(i,j))
      } 
    }
  }
  x_var <- var(Distance[,1])
  y_var <- var(Distance[,2])
  mod <- sqrt(x_var + y_var) 
  return(c(sqrt(x_var), sqrt(y_var), mod))
}

# Calculate p-value via mod of variance of coordinate

prob <- function(x,y){
  p <- length(which(y < x)) / length(y)
  if(p > 0.5){
    return(1 - p)
  }else{
    return(p)
  }
}
```

## Two dimensional plot

```{r}
point1 <- replicate(10000,Coordinate(Permutate_grid(Grid1)))
Grid1_var <- Coordinate(Grid1)

point2 <- replicate(10000,Coordinate(Permutate_grid(Grid2)))
Grid2_var <- Coordinate(Grid2)

point3 <- replicate(10000,Coordinate(Permutate_grid(Grid3)))
Grid3_var <- Coordinate(Grid3)

par(mfrow = c(1,3))

plot(x = point1[1,], y = point1[2,], main = "Two dimentional plot of grid 1", xlab = "x var", ylab = "y var")
segments(x0 = Grid1_var[1], y0 = Grid1_var[2], x1 = Grid1_var[1], y1 = 100, col = "red", lwd = 3)
segments(x0 = Grid1_var[1], y0 = Grid1_var[2], y1 = Grid1_var[2], x1 = 100, col = "red", lwd = 3)

plot(x = point2[1,], y = point2[2,], main = "Two dimentional plot of grid 2", xlab = "x var", ylab = "y var")
segments(x0 = Grid2_var[1], y0 = Grid2_var[2], x1 = Grid2_var[1], y1 = 100, col = "red", lwd = 3)
segments(x0 = Grid2_var[1], y0 = Grid2_var[2], y1 = Grid2_var[2], x1 = 100, col = "red", lwd = 3)

plot(x = point3[1,], y = point1[2,], main = "Two dimentional plot of grid 3", xlab = "x var", ylab = "y var")
segments(x0 = Grid3_var[1], y0 = Grid3_var[2], x1 = Grid3_var[1], y1 = 100, col = "red", lwd = 3)
segments(x0 = Grid3_var[1], y0 = Grid3_var[2], y1 = Grid3_var[2], x1 = 100, col = "red", lwd = 3)
```

## Histogram and p-value
In order to calculate p-value, we can use mode of point the evaluate the distribution.
```{r}
p1 <- prob(Grid1_var[3], point1[3,])
hist(point1[3,], main = paste("Histogram of Grid1", "\n", "p =", p1), xlab = "mode")
abline(v = Grid1_var[3], lwd = 3, col = "red")



p2 <- prob(Grid2_var[3], point2[3,])
hist(point2[3,], main = paste("Histogram of Grid2", "\n", "p =", p2), xlim = c(Grid2_var[3],11), xlab = "mode")
abline(v = Grid2_var[3], lwd = 3, col = "red")


p3 <- prob(Grid3_var[3], point3[3,])
hist(point3[3,], main = paste("Histogram of Grid3", "\n", "p =", p3), xlab = "mode")
abline(v = Grid3_var[3], lwd = 3, col = "red")
```

## Discussion

If we set $\alpha$ = 0.05, then grid1 is statistically signifcantly homogenuously distributed;

If we set $\alpha$ = 0.05, then grid1 is not statistically signifcantly homogenuously distributed (p = 0.09).

Grid2 is absolutely not statistically signifcantly homogenuously distributed (p < 0.00001).

Grid3 is tatistically signifcantly homogenuously distributed (p = 0.4).