---
title: "Expotential task 1"
author: "Chao XIA"
date: "01/14/2019"
output: 
  pdf_document:
    includes:
      in_header: header.tex
    keep_tex: yes
    latex_engine: xelatex
---
## Generate X1 with exponential distribution larger than y

```{r}
set.seed(2018)
X1<-numeric()
GenerateX1<-function(y,t){
  x<-0
  while (x<y){
    x <- rexp(1,t)
  }
  return(x)
}
n = 1000
for (i in 1:n){
  X1[i] = GenerateX1(1,1)
}
```

## Plot histogram and Density Curve

```{r}
X2 <- rexp(1000,1)
output <- X1-1

par(mfrow = c(1,2))
h <- hist(output, main = 'Output', xlab = 'X1-1', ylim = 0:1, freq = F, breaks = 10)
par(new = T)
curve(dexp(x), col = 'red', lwd = 2, xlab = '', ylab = '', xlim = range(h$breaks), ylim = 0:1)
hist(X2, main = 'Contrast', freq = F, breaks = 10, ylim = 0:1)
```
## Discussion

From the these two figure, we found that the distribution of exponential conditional on y and t is just as same as the exponential density function.

I think this is because the memoryless property of Expotential Distribution, which means:

$$
P(T>s+t|T>t) = P(T>s)
$$
In this example,
$$\begin{split}
&P(T>0+1|T>1) = P(T>0) \\
&P(T>1+1|T>1) = P(T>1) \\
&and\ so\ on
\end{split}$$
it is obvious that their distribution will be the same.
