---
title: "The Polya Urn version 2"
author: "Chao XIA"
date: "02/10/2019"
output: pdf_document
---

## The Polya Urn version 2
a. Add a black (special color) ball. Start with two balls, one of which is black.
b. If you draw a non-black ball, proceed as before (add an additional ball of the same color)。
c. If you draw a black ball do the following:
    1. Keep the black ball for now.
    2. Draw another ball from the urn.
    3. Change the color of the second ball, making it a color that is not currently in the urn.
    4. Put both balls back in the urn.

## My code
```{r}
set.seed(1992)

n <- 50 # the final ball amount without black ball

Urn_v2 <- function(finalball, starting_ball){
  
  ball <- starting_ball # define the starting cituation
  
  while(length(ball) < finalball){
    
    pick <- sample(ball, 1) # pick one ball from the urn
    
    if (pick == "0"){ # 0 is black
      
      ball <- ball[-which(ball == pick)] # pick the black ball
      
      pick2 <- sample(1:length(ball), 1) # pick a ball from the remaining balls
      
      ball[pick2] <- as.character(length(ball) + 1) # change the ball's color (since each iteration the length of ball will increase or remain the same, we'll never repeat the ball's color)
      
      ball <- c(ball, pick) # put back the black ball
      
    }else{
      
      ball <- c(ball, pick) # else just add a ball with the picked color
      
    }
    
  }
  
  return(ball)
  
}

end <- numeric() # store the number of (non-black) colors at the end
common <- numeric() # store the number of balls of the commonest color at the end
repeattimes <- 10000
for (i in 1:repeattimes) {
  result <- Urn_v2(n+1, c("1","0"))
  end[i] <- length(table(result)) - 1
  common[i] <- max(table(result))
}

hist(end)
hist(common)
```
```{r}
for (i in 1:repeattimes) {
  result <- Urn_v2(n+1, c("1","1","0"))
  end[i] <- length(table(result)) - 1
  common[i] <- max(table(result))
}

hist(end)
hist(common)
```
